\name: A data-driven few-process model for nucleosynthesis

### 2023-03-29 Paper/main.tex
Updated text on fiducial model

% Abundance and stellar macros
\newcommand{\mgfe}[0]{[{\rm Mg/Fe}]} 
\newcommand{\Acc}{A_{\rm cc}}
\newcommand{\AIa}{A_{\rm Ia}} 
\newcommand{\RIa}{R_{\rm Ia}^X}
\newcommand{\aIa}{\alpha_{\rm Ia}}
\newcommand{\acc}{\alpha_{\rm cc}}
\newcommand{\afe}[0]{[\alpha/{\rm Fe}]}
\newcommand{\femg}{[{\rm Fe}/{\rm Mg}]} 
\newcommand{\xmg}{[{\rm X}/{\rm Mg}]} 
\newcommand{\mgh}{[{\rm Mg}/{\rm H}]}
\newcommand{\feh}[0]{[{\rm Fe/H}]} 
\newcommand{\pIa}{p_{\rm Ia}^{\rm X}}
\newcommand{\pcc}{p_{\rm cc}^{\rm X}}
\newcommand{\fcc}{f_{\rm cc}}
\newcommand{\xfe}{[{\rm X}/{\rm Fe}]} 
\newcommand{\pIasun}{p_{\rm Ia, \odot}^{\rm X}}
\newcommand{\pccsun}{p_{\rm cc, \odot}^{\rm X}}
\newcommand{\logg}{\log(g)}
\newcommand{\teff}{T_{\rm eff}}
\newcommand{\kpc}{\rm \; kpc}
\newcommand{\kel}{\rm \; K}
\newcommand{\msun}{M_{\odot}}
\newcommand{\rgc}{R_{\rm GC}}
\newcommand{\msunvice}{M_{\odot}/M_{\odot \rm formed}}
\newcommand{\MZAMS}{M_{\rm ZAMS}}
\newcommand{\Msun}{M_{\odot}}
\newcommand{\qIax}{q_{\rm Ia}^{\rm X}}
\newcommand{\qccx}{q_{\rm cc}^{\rm X}}
\newcommand{\qagbx}{q_{\rm AGB}^{\rm X}}
\newcommand{\qIa}{q_{\rm Ia}}
\newcommand{\qcc}{q_{\rm cc}}
\newcommand{\qagb}{q_{\rm AGB}}
\newcommand{\Aagb}{A_{\rm AGB}}
\newcommand{\eumg}{[{\rm Eu}/{\rm Mg}]} 
\newcommand{\mgmn}{[{\rm Mg}/{\rm Mn}]} 
\newcommand{\alfe}{[{\rm Al}/{\rm Fe}]} 
\newcommand{\omg}{[{\rm O}/{\rm Mg}]} 
\newcommand{\xo}{[{\rm X}/{\rm O}]} 
\newcommand{\xh}{[{\rm X}/{\rm H}]} 
\newcommand{\angstrom}{\textup{\AA}}
\newcommand{\ejg}[1]{\textcolor{red}{EJG: #1}}

@@

\begin{abstract}\noindent % trust Hogg
Stellar surface abundances look like they are produced by two dominant processes, one prompt and one delayed.
We analyze YYY abundance ratios (ratioed to Mg) for XXXX stars from APOGEE DR17 with a flexible two-process model, in which all element process amplitudes are free parameters.
The two-process model describes each element abundance of each star as the vector sum of a CCSN and SNIa process.
Prior work derives process amplitudes from median observed high-Ia and low-Ia abundance trends; we relax these assumptions and simultaneously fit all model parameters to the data, with only minimal constraints to keep the processes interpretable.
We compare fit parameters from this new method to those derived from prior work [that the two methods produce similar results but with XXX differences and identify similar outlier stars??].
Our flexible method---dubbed \name---tests the validity of the standard two-process assumptions, such as Mg being a pure CCSN element, and that Solar abundances are generated by a roughly half-half combination of CCSN and SNIa processes.
We find [that the assumptions of prior work are generally confirmed. But that the inferred Fcc value of Fe peak elements is strongly dependent upon the chosen CCSN process strength for Fe].
We capitalize on the flexibility to explore the addition of a third process, agnostic about its origin.
The third process most improves [X elements? most significant for X elements?].
Most importantly, \name{} can be extended to populations at very low metallicity, and populations without two clear abundance sequences, such as the LMC/SMC, Sagittarius, and Gaia-Enceladus (??), making it a general diagnostic for enrichment; in principle, it can also be used to constrain fundamental nucleosynthetic parameters.
\end{abstract}

@@

\section{Data Samples}\label{sec:data}

In this paper, we employ stellar abundances from APOGEE DR17 \citep{abdurrouf2022}, part of SDSS-IV \citep{majewski2017}. The APOGEE survey obtains high-resolution ($R\sim22,500$) near-infrared (IR) observations \citep{wilson2019} for stars in the Galactic disk, halo, bulge, and nearby satellites/streams. Observations are taken with two nearly identical spectrographs on the 2.5m Sloan Foundation telescope \citep{wilson2019} at Apache Point Observatory in New Mexico and the 2.5m du Pont Telescope \citep{bowen1973} at the Las Companas Observatory in Chile. Spectral data are reduced and calibrated with the APOGEE data processing pipeline \citep{nidever2015}, after which stellar parameters and abundances are calculated with ASPCAP \citep[APOGEE Stellar Parameter and Chemical Abundance Pipeline][]{holtzman2015, garcia2016}. See \citet[][DR16]{jonsson2020} and Holtzman et al. (in prep., DR17) for a more detailed description of APOGEE data reduction and analysis, and \citet{zasowski2013, zasowski2017} and \citet{santana2021} for a discussion of survey targeting.

APOGEE DR17 reports stellar parameters, including $\teff$ and $\logg$, as well as 20 elemental abundances: C, \ion{C}{1}, N, O, Na, Mg, Al, Si, S, K, Ca, \ion{Ti}{1}, \ion{Ti}{2}, V, Cr, Mn, Fe, Co, Ni, and Ce for 657,135 stars. Among these elements and ions, some are measured more precisely than others. We exclude Ti from our analysis as there are large differences between the abundances derived from the \ion{Ti}{1} and \ion{Ti}{2} lines \citep{jonsson2020}. We also exclude P, as the P abundances are measured from a few very weak spectra features, and the abundance distribution displays strong artifacts \citep{jonsson2020}. Among the remaining elements, we note the following concerns: weak Na spectral features, large abundances scatter in S, significant systematic artifacts in Cr abundances at super-solar metallicities, strong unaccounted for NLTE effects on Mn abundances, and large abundance scatter in Co and Ce--both derived from one line. Further, we note that DR17 is the first time Ce abundances have been included in APOGEE data products. For a more detailed discussion of abundance systematics and their effects on population trends, see \citet{jonsson2020} and \citet{griffith2021a}.

% \begin{itemize}
% \itemsep0em
%     \item Na - Spectral features are weak, one of the least precise elements in APOGEE (but matches optical data pretty well)
%     \item S - Large abundance scatter in observations and no verification with optical data
%     \item Cr - Significant systematic artifacts at super solar metallicity \citep{griffith2021a}
%     \item Mn - Strong NLTE effects exist that are not taken into account
%     \item Co - Abundance derived from one line, causing larger abundance scatter
%     \item Ce - Abundance derived from one line. DR17 is the first time Ce abundances have been included in APOGEE data products. 
%     \item Al,  V, and Mn require large offsets to place solar metallicity stars at solar [X/Fe].
% \end{itemize}

In this paper, we will explore two subsamples of APOGEE DR17 data: (1) the W22 sample, optimized to minimize statistical and systematic errors while probing a large portion of the Galactic disk, and (2) a sample with less stringent cuts, but better coverage of the low-metallicity ([Fe/H] $< -0.5$) disk. We discuss the model fits to the W22 sample in Section~\ref{sec:optimize} and ~\ref{sec:results_W22} and the fits the expanded sample in Section~\ref{sec:results_broad}.

\subsection{W22 Sample}

W22 select a subset of APOGEE DR17 stars with the goal of minimizing statistical errors from poor observations and systematic errors from abundance trends with $\teff$ and/or $\logg$ while preserving a sufficient number of stars to conduct a meaningful statistical analysis across the Galactic disk. To remove poor quality data points, W22 require ASPCAP flags \texttt{STAR\_BAD} \texttt{NO\_ASPCAP\_RESULT} equal zero and remove all stars with \texttt{FE\_H\_FLAG} or \texttt{MG\_FE\_FLAG} set. They only include stars from the main survey sample (\texttt{EXTRATARG} = 0), and use named abundances (\texttt{X\_FE}), as recommended by \citet{jonsson2020}. In addition to these quality cuts, W22 apply the following sample selection:
\begin{itemize}
\itemsep0em
    \item $R=3-13$ kpc, $|Z| \leq 2$ kpc (distances from \citealp{leung2019})
    \item $-0.75 \leq \mgh \leq 0.45$
    \item S/N $\geq 200$ for $\mgh > -0.5$ and S/N $\geq 100$ for $\mgh < -0.5$
    \item $\logg = 1-2.5$ dex
    \item $\teff = 4000-4600$ K
\end{itemize}
By selecting stars from a small range of $\logg$ and $\teff$, W22 eliminate the red clump stars \citep{vincenzo2021a} and minimize stellar parameter-dependent abundance trends \citep[see][]{griffith2021a}. The cuts result in a sample of 34,410 stars.

W22 present abundance for Mg, O, Si, S, Ca, C+N, Na, Al, K, Cr, Fe, Ni, V, Mn, Co, and Ce. In the analysis of each element, X, W22 drop stars with \texttt{X\_FE\_FLAG} (maximum of $\sim 560$ for Ce). While the surface abundances of C and N differ from the stellar birth abundances for RGB stars due to the CNO processes and dredge up events \citep{iben1965, shetrone2019}, the total C+N abundance remains constant. W22 consider C+N as an element in their analysis, taking [(C+N)/H] to be 
\begin{equation}
    [\text{C+N}/\text{H}] = \log_{10}(10^{\text{[C/H]}+8.39} + 10^{\text{[N/H]}+7.78}) - \log_{10}(10^{8.39} + 10^{7.78})
\end{equation}
using logarithmic solar abundances for C (8.39) and N (7.78) from \citet{grevesse2007}.

%\ejg{W22 use X/Fe error from APOGEE and don't convert it to X/Mg....Use C/Fe error for the C+N value because the ratio is mostly C... Need error discussion here and in next section.}

In this paper, we will analyze the same set of stars and elements as W22 (denoted ``W22 sample''), adopting the same definition of [C+N/H]. We plot the distribution of abundances in [X/Mg] vs. [Mg/H] for this sample in Figure~\ref{fig:w22_xmg}. All elemental abundances have zero-point offsets applied and have been de-trended for correlations with $\teff$, as in W22 (see their Table 1).

\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{Figures/W22_xmg.pdf}
    \caption{Stellar abundance distributions for the W22 sample. The number in the top right corner indicates the number of stars used in the analysis of that element. Elements are ordered by atomic number.}
    \label{fig:w22_xmg}
\end{figure*}

\subsection{Expanded Sample}

While the W22 sample presented above is a good first test case of \name, we are interested in the behavior of our model when applied to a wider sample of data, in particular extending the metalicity range to lower values of [Mg/H]. To construct this sample, we apply the same quality cuts as taken by W22 (\texttt{STAR\_BAD} = 0, \texttt{NO\_ASPCAP\_RESULT} = 0, \texttt{FE\_H\_FLAG} = 0, \texttt{MG\_FE\_FLAG} = 0, and \texttt{EXTRATARG} = 0). We add a cut for stars with the \texttt{ROTATION\_WARNING} flag set. We additionally take the following sample selection cuts: 
\itemsep0em
\begin{itemize}
    \item S/N $\geq 100$
    \item $\logg =  0 - 2.5$ dex
    \item $\teff = 4000-4600$ K
    \item H $< 12$
\end{itemize}
We make no cuts on the [Mg/H] abundances. The wider $\logg$ range expands the sample of giants, though we select stars in the same small $\teff$ range as W22 to exclude the red clump \citep{vincenzo2021a} and cool stars that show abundance artifacts \citep{jonsson2020}. The $H$-band magnitude cut acts as a proxy for a distance cut \ejg{Add how many stars it excludes.}
Our final sample includes 51,386 stars and extends to [Mg/H]=-2.2. As in W22, we only analyze [X/Mg] abundances for stars where the [X/Fe] value is unflagged. This has a minimal effect for most elements but reduces the sample size to 49,851 for Ce.

\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{Figures/xmg.pdf}
    \caption{Stellar abundance distributions for our expanded stellar sample. The number in the top right corner indicates the number of stars used in the analysis of that element. Elements are ordered by atomic number. No zero-point offsets have been applied.}
    \label{fig:exp_xmg}
\end{figure*}

We plot the distribution of abundances in [X/Mg] vs. [Mg/H] for the expanded sample in Figure~\ref{fig:exp_xmg}. We do not apply zero-point offsets or temperature calibrations, as done in the W22 sample.

The expansion of the data sample does introduce abundance systematics that the W22 cuts sought to exclude. As discussed in \citet{jonsson2020} and \citet{griffith2021a}, APGOEE DR16 and DR17 suffer from poorly understood abundance artifacts, such as a ``finger'' in [$\alpha$/M] and low-[X/Fe] banding in elements such as Al and Ni. In our expanded data set, we observe potential abundance artifacts in Al, K, V, and Cr, though the strict temperature cuts help to minimize these effects.

\section{The K-Proccess Model}\label{sec:model}

@@

where there are $i$ stars and $j$ elements described by process amplitudes $A_i^{\rm CC}$ and $A_i^{\rm Ia}$ and process vectors $\vec{q}_{{\rm CC}, j}^{\,Z}$ and $\vec{q}_{{\rm Ia}, j}^{\,Z}$. The observed value of $\xh$ can be described as the model plus observational noise and/or contributions from an additional non-CCSN nor SNIa processes:

@@

While such a model could be fit to a set stellar abundances, there are many degenerate solutions when there are no constraints on the model parameters. To better define our model and to collapse such degeneracies, we put forth a set a set of assumptions in the following section.

@@

\paragraph{1. Two processes}
The majority of $\alpha$, light odd-Z, and Fe-peak elements are dominantly produced by two sources: CCSN and SNIa. This is substantiated by theoretical yields \citep[e.g.,][]{anderson2019, rybizki2017} and past successful models (e.g., G22, \citealp{ting2022}, W22).
However, the assumption fails to allow for enrichment from other known sources, such as merging neutron stars, unique classes of supernoave, and AGB stars, the latter of which significantly contribute to C+N and Ce production.

@@

\paragraph{3. Non-negativity}
All process amplitudes for all elements are non-negative (elements considered here are only produced, not ever destroyed, by the two processes),
and all stars are made of a sum of the two non-negative processes.
This makes the model similar to a non-negative matrix factorization (\citealt{nnmf} \ejg{CITATION}). This assumption is not enforced in G22 and W22.

\paragraph{4. Mg is a pure CCSN}
All magnesium is produced in the CCSN process and none in the SNIa process. This is substantiated by theoretical yields \citep[e.g.,][]{anderson2019, rybizki2017}.
This breaks a rotational symmetry in the process space and makes the processes interpretable in terms of SN types.

\paragraph{5. Metallicity dependence}
The fact that all Mg is produced in the CCSN process is an assumption that is independent of metallicity.
All other process vectors are permitted to float as a function of metallicity.
The only exceptions are the Mg level in the CCSN process and the Fe level in the SNIa process, but these constraints are just model normalizations; they don't restrict the overall model freedom, which is about abundance ratios. \ejg{Check that this is right. Especially for Fe.}

@@

\paragraph{6. APOGEE abundances and uncertainties}
We assume that the APOGEE abundances and uncertainties can be used for this project.
This is not the same as assuming that they are correct, but rather we are assuming that it is possible and useful to build an interpretable 2-process model to explain them. We describe the potential data systemtics in Section~\ref{sec:data}. The APOGEE errors are likely underestimated for many more challenging elements.

@@

To enforce assumptions 4 and 5, we initialize the model with specific values of $\vec{q}_{{\rm CC}}$ and $\vec{q}_{{\rm Ia}}$ and place regularizations ($\lambda$) on these constraints so that the model disfavors an alternative solution. In the initial model, we strongly require that
\begin{equation}\label{eq:qcc_solar}
    \vec{q}_{{\rm CC, Mg}}^{\,Z_{\odot}} = 1, \quad \vec{q}_{{\rm Ia, Mg}}^{\,Z_{\odot}} = 0, \quad 
    \vec{q}_{{\rm CC, Fe}}^{\,Z_{\odot}} = 0.5, \quad \vec{q}_{{\rm Ia, Fe}}^{\,Z_{\odot}} = 0.5
\end{equation}
at solar metallicity and

\begin{equation}\label{eq:qcc_z}
    \vec{q}_{{\rm CC, Mg}}^{\,Z} = \vec{q}_{{\rm CC, Mg}}^{\,Z_{\odot}}, \quad 
    \vec{q}_{{\rm CC, Fe}}^{\,Z} = \vec{q}_{{\rm CC, Fe}}^{\,Z_{\odot}},  \quad 
    \vec{q}_{{\rm Ia, Fe}}^{\,Z} = \vec{q}_{{\rm Ia, Fe}}^{\,Z_{\odot}}
\end{equation}
at all other metallicities.
\ejg{Do we want to list regularization strengths? What is the difference between lambda a and lambda b in the notebook?} 
While the CCSN process constraint (that Mg is a pure CCSN element) is grounded in nucleosynthetic theory, there is no equivalent nucleosynthetic fact to constrain the SNIa process. We begin by initializing our model with the assumption that $\vec{q}_{{\rm CC, Fe}} = 0.5$, as in G22 and W22. This assumption places a pure CCSN star on the low metallicity [Mg/Fe] plateau near 0.3 dex, in agreement with APOGEE observations, but in contention with recent results from \citet{conroy2022} which places the [Mg/Fe] plateau near 0.6 dex. The flexibility of \name allows us to fit the model with different values of $\vec{q}_{{\rm CC, Fe}}$, allow for metallicity dependence in Fe CCSN and SNIa yields, and relax the regularization strengths in our initialization to see if the model prefers a different process vector. We explore different choices of $\vec{q}_{{\rm CC, Fe}}$ and regularization strengths in Section~\ref{subsec:fcc}.

We initialize the model according to Equations~\ref{eq:qcc_solar} and~\ref{eq:qcc_z}, setting all other $\vec{q}_{{\rm CC,}j}$ = $\vec{q}_{{\rm Ia,}j} = 0.5$ and all $A_i^{\rm CC} = A_j^{\rm Ia} = 0$. \ejg{Is this right?}

We then fit the $K$-process model to the W22 sample by minimizing the a chi-squared ($\chi^2$) with Gauss-Newton nonlinear least-squares solver in \texttt{jaxopt} where
\begin{equation}
    \chi^2 = \sum_{i=1}^{N} (\xh_{ij} - m_{ij})^{\rm T} c_j^{-1} (\xh_{ij} - m_{ij}) + \lambda~,
\end{equation}
and $c_j^{-1}$ is a matrix of the inverse variance squared,
\begin{equation}
    c^{-1} = \begin{bmatrix}
\frac{1}{\sigma_j^2} & 0 & ...\\
0 & \frac{1}{\sigma_j^2} & ... \\
0 & 0 & ... 
\end{bmatrix} .
\end{equation}
We fit the model to $N$ stars, such that $1 \leq i \leq N$, and $M$ elements, such that $1 \leq j \leq M$ where $j \in \{ \text{Mg, O, Si, S, Ca, CN, Na, Al, K, Cr, Fe, Ni, V, Mn, Co, Ce} \}$, by iteratively optimizing the process vectors (dubbed the q-step) and the process amplitudes (dubbed the A-step). \ejg{Add bit about how we determine how many times to repeat the iterations.}

\section{Optimization of \name} \label{sec:optimize}

\section{Comparing the W22 Two-Process Model and \name} \label{sec:results_W22}

\begin{itemize}
    \item Differences/similarities between the W22 fits and our fits. What do the differences mean
    \item Dependence of results on assumptions on assumptions such as the Fe yield/plateau value. Contrast the results for different assumptions and show the uncertainty in fcc 
\end{itemize}

\subsection{Dependence of $\fcc$ on Assumptions} \label{subsec:fcc}

>>>>

% Abundance and stellar macros
\newcommand{\mgfe}[0]{[{\rm Mg/Fe}]} 
\newcommand{\afe}[0]{[\alpha/{\rm Fe}]}
\newcommand{\femg}{[{\rm Fe}/{\rm Mg}]} 
\newcommand{\xmg}{[{\rm X}/{\rm Mg}]} 
\newcommand{\mgh}{[{\rm Mg}/{\rm H}]}
\newcommand{\feh}[0]{[{\rm Fe/H}]} 
\newcommand{\xfe}{[{\rm X}/{\rm Fe}]} 
\newcommand{\xh}{[{\rm X}/{\rm H}]}

\newcommand{\logg}{\log(g)}
\newcommand{\teff}{T_{\rm eff}}
\newcommand{\kpc}{\rm \; kpc}
\newcommand{\kel}{\rm \; K}
\newcommand{\msun}{M_{\odot}}
\newcommand{\rgc}{R_{\rm GC}}
\newcommand{\msunvice}{M_{\odot}/M_{\odot \rm formed}}
\newcommand{\MZAMS}{M_{\rm ZAMS}}

\newcommand{\qcc}{\vec{q}_{{\rm CC,}j}^{Z}}
\newcommand{\qccFe}{\vec{q}_{{\rm CC, Fe}}^{Z}}
\newcommand{\dqccFe}{d\vec{q}_{{\rm CC, Fe}}^{Z}/dZ}
\newcommand{\qIa}{\vec{q}_{{\rm Ia,}j}^{Z}}
\newcommand{\qIaFe}{\vec{q}_{{\rm Ia,Fe}}^{Z}}
\newcommand{\Acc}{A^{\rm CC}_{i}}
\newcommand{\AIa}{A^{\rm Ia}_{i}}
\newcommand{\fcc}{f_{\rm CC}}

\newcommand{\ejg}[1]{\textcolor{red}{EJG: #1}}

@@

\begin{abstract}\noindent % trust Hogg

\ejg{Need to update abstract}

Stellar surface abundances look like they are produced by two dominant processes, one prompt and one delayed.
We analyze YYY abundance ratios (ratioed to Mg) for XXXX stars from APOGEE DR17 with a flexible two-process model, in which all element process amplitudes are free parameters.
The two-process model describes each element abundance of each star as the vector sum of a CCSN and SNIa process.
Prior work derives process amplitudes from median observed high-Ia and low-Ia abundance trends; we relax these assumptions and simultaneously fit all model parameters to the data, with only minimal constraints to keep the processes interpretable.
We compare fit parameters from this new method to those derived from prior work [that the two methods produce similar results but with XXX differences and identify similar outlier stars??].
Our flexible method---dubbed \name---tests the validity of the standard two-process assumptions, such as Mg being a pure CCSN element, and that Solar abundances are generated by a roughly half-half combination of CCSN and SNIa processes.
We find [that the assumptions of prior work are generally confirmed. But that the inferred Fcc value of Fe peak elements is strongly dependent upon the chosen CCSN process strength for Fe].
We capitalize on the flexibility to explore the addition of a third process, agnostic about its origin.
The third process most improves [X elements? most significant for X elements?].
Most importantly, \name{} can be extended to populations at very low metallicity, and populations without two clear abundance sequences, such as the LMC/SMC, Sagittarius, and Gaia-Enceladus (??), making it a general diagnostic for enrichment; in principle, it can also be used to constrain fundamental nucleosynthetic parameters.
\end{abstract}

@@

\section{The K-Process Model}\label{sec:model}

@@

where there are $i$ stars and $j$ elements described by process amplitudes $A_i^{\rm CC}$ and $A_i^{\rm Ia}$ and process vectors $\vec{q}_{{\rm CC}, j}^{\,Z}$ and $\vec{q}_{{\rm Ia}, j}^{\,Z}$. The observed value of [X/H] can be described as the model plus observational noise and/or contributions from an additional non-CCSN nor SNIa processes:

@@

While such a model could be fit to a set of stellar abundances, there are many degenerate solutions when there are no constraints on the model parameters. To better define our model and to collapse such degeneracies, we put forth a set of assumptions in the following section.

@@

\paragraph{1. Two processes}
The majority of $\alpha$, light odd-Z, and Fe-peak elements are dominantly produced by two sources: prompt CCSN and delayed SNIa. This is substantiated by theoretical yields \citep[e.g.,][]{anderson2019, rybizki2017} and past successful models (e.g., G22, \citealp{ting2022}, W22).
However, the assumption fails to allow for enrichment from other known sources, such as merging neutron stars, unique classes of supernoave, and AGB stars, the latter of which contribute to C+N and Ce production.

@@

\paragraph{3. Non-negativity}
All process amplitudes for all elements are non-negative (elements considered here are only produced, not ever destroyed, by the two processes),
and all stars are made of a sum of the two non-negative processes.
This makes the model similar to a non-negative matrix factorization (\citealt{nnmf} \ejg{CITATION}). This assumption is not enforced in G22 and W22 where both process vectors and amplitudes can be negative.

\paragraph{4. Mg is a pure CCSN element}
All magnesium is produced in the CCSN process and none in the SNIa process, substantiated by theoretical yields \citep[e.g.,][]{anderson2019, rybizki2017} This breaks a rotational symmetry in the process space and makes the processes interpretable in terms of SN types.

\paragraph{5. Metallicity dependence}
The fact that all Mg is produced in the CCSN process is an assumption that is independent of metallicity.
All other process vectors, except Fe, are permitted to float as a function of metallicity.
We further fix the Fe CCSN and SNIa processes, though vary the Fe assumptions as discussed in Section~\ref{subsec:qcc_fe}.
These constraints are just model normalizations; they don't restrict the overall model freedom.

@@

\paragraph{6. APOGEE abundances and uncertainties}
We assume that the APOGEE abundances and uncertainties can be used for this project.
This is not the same as assuming that they are correct, but rather we are assuming that it is possible and useful to build an interpretable Few-Process model to explain them. We describe the potential data systematics in Section~\ref{sec:data}. The APOGEE errors are likely underestimated for many elements.

@@

To enforce assumptions 4 and 5, we initialize the model with specific values of $\vec{q}_{{\rm CC}}$ and $\vec{q}_{{\rm Ia}}$ and place regularizations ($\lambda$) on these constraints so that the model disfavors an alternative solution. In the fiducial model, we strongly require that
\begin{equation}\label{eq:qcc_solar}
    \vec{q}_{{\rm CC, Mg}}^{\,Z_{\odot}} = 1, \quad \vec{q}_{{\rm Ia, Mg}}^{\,Z_{\odot}} = 0, \quad 
    \vec{q}_{{\rm CC, Fe}}^{\,Z_{\odot}} = 0.4, \quad \vec{q}_{{\rm Ia, Fe}}^{\,Z_{\odot}} = 0.6
\end{equation}
at solar metallicity and

\begin{equation}\label{eq:qcc_z}
    \vec{q}_{{\rm CC, Mg}}^{\,Z} = \vec{q}_{{\rm CC, Mg}}^{\,Z_{\odot}}, \quad 
    \vec{q}_{{\rm CC, Fe}}^{\,Z} = \vec{q}_{{\rm CC, Fe}}^{\,Z_{\odot}},  \quad 
    \vec{q}_{{\rm Ia, Fe}}^{\,Z} = \vec{q}_{{\rm Ia, Fe}}^{\,Z_{\odot}}
\end{equation}
at all other metallicities.
\ejg{Do we want to list regularization strengths? What is the difference between lambda a and lambda b in the notebook?} 
While the CCSN process constraint (that Mg is a pure CCSN element) is grounded in nucleosynthetic theory, there is no equivalent nucleosynthetic fact to constrain the SNIa process. We begin by initializing our model with the assumption that $\vec{q}_{{\rm CC, Fe}} = 0.4$. This assumption places a pure CCSN star on the low metallicity [Mg/Fe] plateau near 0.4 dex, in agreement with APOGEE observations, but in contention with recent results from \citet{conroy2022} which places the [Mg/Fe] plateau near 0.6 dex. The flexibility of the Few-Process model allows us to fit the model with different values of $\vec{q}_{{\rm CC, Fe}}$, allow for metallicity dependence in Fe CCSN and SNIa yields, and relax the regularization strengths in our initialization to see if the model prefers a different process vector. We explore different choices of $\vec{q}_{{\rm CC, Fe}}$ and regularization strengths in Section~\ref{subsec:qcc_fe}.

We initialize the model according to Equations~\ref{eq:qcc_solar} and~\ref{eq:qcc_z}, setting all other $\vec{q}_{{\rm CC,}j}$ = $\vec{q}_{{\rm Ia,}j} = 0.5$ and all $A_i^{\rm CC} = A_j^{\rm Ia} = 0$. \ejg{Is this right?}. We then fit the $K$-process model to the W22 sample by minimizing the a chi-squared ($\chi^2$) with Gauss-Newton nonlinear least-squares solver in \texttt{jaxopt} where
\begin{equation}
    \chi^2 = \sum_{i=1}^{N} (\xh_{ij} - m_{ij})^{\rm T} c_j^{-1} (\xh_{ij} - m_{ij}) + \lambda~,
\end{equation}
and $c_j^{-1}$ is a matrix of the inverse variance squared,
\begin{equation}
    c^{-1} = \begin{bmatrix}
\frac{1}{\sigma_j^2} & 0 & ...\\
0 & \frac{1}{\sigma_j^2} & ... \\
0 & 0 & ... 
\end{bmatrix} .
\end{equation}
We fit the model to $N$ stars, such that $1 \leq i \leq N$, and $M$ elements, such that $1 \leq j \leq M$ where $j \in \{ \text{Mg, O, Si, S, Ca, C+N, Na, Al, K, Cr, Fe, Ni, Mn, Co, Ce} \}$, by iteratively optimizing the process vectors (dubbed the q-step) and the process amplitudes (dubbed the A-step). In the initialization, we use only Mg and Fe in the A-step. We then repeat the A-step and q-step with all elements for all stars 48 times, only updating the best-fit process parameters when the objective function improves. We find few differences in the model when we decrease the number of iterations to 32, indicating that the model quickly finds a good solution.

In the few-process model, all stars and all unflagged abundances contribute to the optimization of $\qcc$, $\qIa$, $\Acc$, and $\AIa$. Outlier stars with large abundance errors will contribute minimally to the $\chi^2$ minimization, but outlier stars with truly unique abundances (and small errors) will pull the model away from the values favored by the larger population. To mitigate this effect, we artificially inflate the errors on abundances with values of $\chi^2$ such that 
\begin{equation}\label{eq:inflate_ivar}
    \frac{1}{\sigma^2} = \frac{Q^2/\sigma_{\rm obs}^2}{Q^2 + (\xh_{\rm obs} - \xh_{\rm pred})^2 /\sigma_{\rm obs}^2 },    
\end{equation}
where $Q$ is the softening parameter. We inflate the errors after the initialization (with $Q=7$) and after the 16th and 32nd iteration (with $Q=5$).

\ejg{DH should check all of this and add some more statistics, I'm sure I have explained something incorrectly.}

\section{Data}\label{sec:data}

In this paper, we employ stellar abundances from APOGEE DR17 \citep{abdurrouf2022}, part of the SDSS-IV \citep{majewski2017}. The APOGEE survey obtains high-resolution ($R\sim22,500$) near-infrared (IR) observations \citep{wilson2019} for stars in the Galactic disk, halo, bulge, and nearby satellites/streams. Observations are taken with two nearly identical spectrographs on the 2.5m Sloan Foundation telescope \citep{wilson2019} at Apache Point Observatory in New Mexico and the 2.5m du Pont Telescope \citep{bowen1973} at the Las Companas Observatory in Chile. Spectral data are reduced and calibrated with the APOGEE data processing pipeline \citep{nidever2015}, after which stellar parameters and abundances are calculated with ASPCAP \citep[APOGEE Stellar Parameter and Chemical Abundance Pipeline][]{holtzman2015, garcia2016}. See \citet[][DR16]{jonsson2020} and Holtzman et al. (in prep., DR17) for a more detailed description of APOGEE data reduction and analysis, and \citet{zasowski2013, zasowski2017} and \citet{santana2021} for a discussion of survey targeting.

APOGEE DR17 reports stellar parameters, including $\teff$ and $\logg$, as well as 20 elemental abundances: C, \ion{C}{1}, N, O, Na, Mg, Al, Si, S, K, Ca, \ion{Ti}{1}, \ion{Ti}{2}, V, Cr, Mn, Fe, Co, Ni, and Ce for 657,135 stars. Among these elements and ions, some are measured more precisely than others. We exclude Ti from our analysis as there are large differences between the abundances derived from the \ion{Ti}{1} and \ion{Ti}{2} lines \citep{jonsson2020}. We also exclude P and V, as the P abundances are measured from a few very weak spectra features and V abundances are one of the least precise and least accurate labels \citep{jonsson2020}, Both P and V display strong abundance artifacts and large scatter. Among the remaining elements, we note the following concerns: weak Na spectral features, large abundances scatter in S, significant systematic artifacts in Cr abundances at super-solar metallicities, strong unaccounted for NLTE effects on Mn abundances, and large abundance scatter in Co and Ce--both derived from one line. Further, we note that DR17 is the first time Ce abundances have been included in APOGEE data products. For a more detailed discussion of abundance systematics and their effects on population trends, see \citet{jonsson2020} and \citet{griffith2021a}.

In this paper, we will explore two subsamples of APOGEE DR17 data: (1) a model training set, similar to the W22 sample, that is optimized to minimize statistical and systematic errors while probing a large portion of the Galactic disk, and (2) a testing sample comprised of all unflagged DR17 giant stars above [Fe/H] of $-0.75$. We discuss the training sample in depth in Sections~\ref{sec:fidutial} and ~\ref{sec:variations} and the fits to the testing sample in Section~\ref{sec:results_broad}.

\subsection{Training Sample}\label{subsec:train}

For our training sample, we select a subset of APOGEE DR17 stars with the goal of minimizing statistical errors from poor observations and systematic errors from abundance trends with $\teff$ and/or $\logg$ while preserving a sufficient number of stars to conduct a meaningful statistical analysis across the Galactic disk. To remove poor quality data points, we require ASPCAP flags \texttt{STAR\_BAD} \texttt{NO\_ASPCAP\_RESULT} equal zero. We only include stars from the main survey sample (\texttt{EXTRATARG} = 0), and use named abundances (\texttt{X\_FE}), as recommended by \citet{jonsson2020}. In addition to these quality cuts, we apply the following sample selection:
\begin{itemize}
\itemsep0em
    \item $\mgh > -0.75$
    \item S/N $\geq 200$
    \item $\logg = 1-3$ dex
    \item $\teff = 4000-5200$ K.
\end{itemize}
To eliminate red clump (RC) stars, which show abundance variations from the RGB sample \citep{vincenzo2021a}, we cross-match with and remove stars that appear in the APOGEE RC VAC \ejg{Add Citation}. These cuts result in a sample of 48,659 stars.

% W22 employ a smaller $\logg$ ($1.0-2.5$) and $\teff$ ($4000-4600$) range, add distance requirement ($R=3-13$ kpc with distances from \citealp{leung2019}), and place a weaker SNR requirement ($>100$) for stars with [Mg/H] $< -0.5$. By expanding the W22 sample, we increase the number of stars in the training set by a factor of 1.4.

We present abundance for Mg, O, Si, S, Ca, C+N, Na, Al, K, Cr, Fe, Ni, Mn, Co, and Ce. In the analysis of each element, X, we drop stars with \texttt{X\_FE\_FLAG} (maximum of $\sim 700$ for Ce). While the surface abundances of C and N differ from the stellar birth abundances for RGB stars due to the CNO processes and dredge-up events \citep{iben1965, shetrone2019}, the total C+N abundance remains constant. As in W22, we consider C+N as an element, taking [(C+N)/H] to be 
\begin{equation}
    [\text{C+N}/\text{H}] = \log_{10}(10^{\text{[C/H]}+8.39} + 10^{\text{[N/H]}+7.78}) - \log_{10}(10^{8.39} + 10^{7.78})
\end{equation}
using logarithmic solar abundances for C (8.39) and N (7.78) from \citet{grevesse2007}. We further adopt the error on the [C/Fe] abundance as the error on [C+N/Fe], since C typically dominates in the abundance ratio.
